Including EW and mixed QCD-EW corrections into PDF fits

### 2020-04-01 paper/sec-introduction.tex
Intro - first draft

In the first respect, because Quantum Chromodynamics (QCD) dominates the 
interactions occurring within colliding protons at the LHC, much effort 
has been devoted to the computation of higher-order QCD corrections: 
fully-differential next-to-leading order (NLO) results, possibly matched to the
parton shower, are currently automated in various general-purpose
Monte Carlo generators~\cite{Gleisberg:2008ta,Alwall:2014hca,Bellm:2015jjp}, 
while an increasing number of next-to-next-to-leading order (NNLO) predictions 
are becoming available for processes with various degrees of inclusiveness
(see {\it e.g.}~\cite{Amoroso:2020lgh} and references therein). The computation
of higher-order corrections in the Electroweak (EW) theory has 
witnessed a comparatively less comprehensive progress. Nevertheless frameworks 
were developed in which the QCD and EW couplings are simultaneously treated as 
small parameters in the perturbative expansion, and the computation of 
theoretical predictions, accurate to NLO (including mixed-coupling QCD+EW 
terms), is automated~\cite{Biedermann:2017yoi,Frederix:2018nkq}.

@@

The systematic inclusion of EW corrections in a PDF determination entails
the solution of two separate problems. First, a problem of efficiency:
fast interpolation grids should be constructed, whereby partonic matrix 
elements, accurate to NLO QCD+EW, are precomputed in such a way that the 
numerical convolution with generic input PDFs can be efficiently approximated
by means of interpolation techniques. Such grids are a fundamental ingredient
in any PDF fit, where the evaluation of the hadronic cross section needs to 
be performed a large number of times. While two default formats exist for
these grids, {\sc APPLgrid}~\cite{Carli:2010rw} and 
{\sc fastNLO}~\cite{Kluge:2006xs}, none of them allows for the inlusion of EW
corrections nor for the interface to a Monte Carlo generator accurate to 
NLO QED+EW. Second, a problem of consistency: the way in which EW effects may
(or may not) be folded into the data varies across different experimental 
analyses, a fact that challenges their consistent theoretical interpretation. 
Examples are the subtraction of background processes which should not be 
considered as such ({\it e.g.} the $t$-channel photon-induced component in 
neutral-current Drell-Yan, which are not different processes beyond LO) or of 
just a part of the EW effects ({\it e.g.} multiple-photon radiation from light 
particles in the final state of neutral- or charged-current Drell-Yan, 
especially with electrons). Be that as it may if EW effects are systematically
included in theoretical predictions, they should not be in the data,
otherwise they will be double counted.

In this paper we address the first of these two problems.
How? By developing {\sc PineAPPL}

the second problem, the solution of which, however, remains beyond the scope of 
the current work.

How to do that consistently, include systematically EW corrections, blabla.\\
Outline a possible strategy to deal with EW correction in LHC observables (FSR, PHOTOS, dressed/born, ...).\\
Pinappl kills amcfast and mcgrid, replace applgrid and fastnlo.\\

The paper is organised as follows.

>>>>

In the first respect, because Quantum Chromodynamics (QCD) dominates the 
interactions occurring within colliding protons at the LHC, much effort 
has been devoted to the computation of higher-order QCD corrections: 
fully-differential next-to-leading order (NLO) results, possibly matched to the
parton shower, are currently automated in various general-purpose
Monte Carlo generators~\cite{Gleisberg:2008ta,Alwall:2014hca,Bellm:2015jjp}
(see also ref.~\cite{Buckley:2011ms} for a review),
while an increasing number of next-to-next-to-leading order (NNLO) predictions 
are becoming available for processes with various degrees of inclusiveness
(see {\it e.g.}~\cite{Amoroso:2020lgh} and references therein). The computation
of higher-order corrections in the Electroweak (EW) theory has 
witnessed a comparatively less comprehensive progress. Nevertheless frameworks 
were developed in which the QCD and EW couplings are simultaneously treated as 
small parameters in the perturbative expansion, and the computation of 
theoretical predictions, accurate to NLO (including mixed-coupling QCD+EW 
terms), is automated~\cite{Biedermann:2017yoi,Frederix:2018nkq}.

@@

The systematic inclusion of EW corrections in a PDF determination entails
the solution of two separate problems. First, a problem of efficiency:
fast interpolation grids should be constructed, whereby partonic matrix 
elements, accurate to NLO QCD+EW, are precomputed in such a way that the 
numerical convolution with generic input PDFs can be efficiently approximated
by means of interpolation techniques. Such grids are a fundamental ingredient
in any PDF fit, where the evaluation of the hadronic cross section needs to 
be performed a large number of times. While two default formats exist for
these grids, {\sc APPLgrid}~\cite{Carli:2010rw} and 
{\sc fastNLO}~\cite{Kluge:2006xs}, none of them allows for the inlusion of EW
corrections nor for the interface to a Monte Carlo generator accurate to 
NLO QED+EW. Second, a problem of consistency: the way in which EW effects may
(or may not) be folded into the data varies across different experimental 
analyses, a fact that challenges their consistent theoretical interpretation. 
Examples are the subtraction of background processes which should not be 
considered as such ({\it e.g.} the $t$-channel photon-induced component in 
neutral-current Drell-Yan, which are not different processes beyond LO) or of 
just a part of the EW effects ({\it e.g.} multiple-photon radiation from light 
particles in the final state of neutral- or charged-current Drell-Yan, 
especially with electrons). Be that as it may, if EW effects are systematically
included in theoretical predictions, they should not be included in the data,
otherwise they will be double counted.

In this paper we address the first of these two problems. Specifically, 
we develop {\sc PineAPPL}, a new library that provides a direct interface 
from parton-level Monte Carlo event generators accurate up to NLO QCD+EW to
fast interpolation grids. The library supports variations of the factorisation
and renormalisations scales, resummation, and matching with a photon- and/or
parton-shower. The library is included by default in 
{\sc MadGraph5\_aMC@NLO} ({\sc mg5\_aMC@NLO} henceforth) within which it has 
been designed and tested. However, it is suitable for use as a plugin in any 
Monte Carlo generator that supports NLO QCD+EW computations, such as 
{\sc SHERPA}~\cite{Biedermann:2017yoi}. In this respect, {\sc PineAPPL}
supersedes the {\sc aMCfast} interface~\cite{Bertone:2014zva} and complements
the {\sc MCgrid} interface~\cite{DelDebbio:2013kxa,Bothmann:2015dba}.

The paper is organised as follows. In Sect.~\ref{sec:pineappl} we introduce
{\sc PineAPPL}, we describe its features, we illustrate its operation, and we
assess its performance. In Sect.~\ref{sec:results}, by means of 
{\sc mg5\_AMC@NLO} and {\sc PineAPPL}, we compute fast interpolation grids for 
a representative set of data that is typically included in a fit of PDFs.
In Sect.~\ref{sec:double-counting} we use these results to formulate more 
comprehensively the second problem sketched above, the solution of which, 
however, remains beyond the scope of the current work. We provide an outlook in
Sect~\ref{sec:conclusion}.

%Outline a possible strategy to deal with EW correction in LHC observables 
%(FSR, PHOTOS, dressed/born, ...).
